Combining checkpointing and data compression\\ for high frequency seismic imaging

### 2018-10-03 paper/paper.tex

Our performance model assumes
 that the compression algorithm behaves uniformly
across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
now becomes
\begin{equation}
\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}
  \cdot \mathbf{B}} + t_c) + \mathbf{N}_R \cdot
(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d)
\end{equation}
where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
and $t_d$ are compression and decompression times, respectively. At the same
time, the recomputation overhead decreases
because $\mathbf{F}$ times more checkpoints are now available.

@@

Figure \ref{fig:varying_compute} shows the speedups predicted by the
performance model for kernels of different computational cost. Perhaps
unsurprisingly, the benefit gained by compression drops rapidly if the
time to compute a single timestep is less than the time to compress
the computed data (since recomputing would be cheaper in that case). For
increasing computational cost, the speedup from compression saturates
at a bound. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-compute.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  compute cost. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. The benefits of compression drop rapidly if the
computational cost of the kernel that generated the data is much lower
than the cost of compressing the data. For increasing computational
costs, the benefits are bounded.}
\label{fig:varying_compute}
\end{figure}

Figure \ref{fig:varying_nt} shows the speedups predicted for differing
number of timesteps. It is interesting to note that there is a
consistent increase in the speedup as the number of timesteps goes
up. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-nt.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  number of timesteps to be reversed. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. It can be seen that compression becomes more
beneficial as the number of timesteps is increased.}
\label{fig:varying_nt}
\end{figure}
\section{Conclusions and Future work}
What are the problems?
The performance we developed might be useful in judging whether compression
would be helpful in a particular scenario without actually building a
complete implementation first. We have tried to keep it application
and platform agnostic in order to maximize its applicability. However
there are a few limitations in this model. Firstly, we assumed that
the compression performance we saw for a single checkpoint is
representative of any checkpoint throughout the simulation. This
assumption might have been completely wrong if made about a checkpoint
in the first few time steps since we consistently saw upto 800x
compression in those cases - since they are mostly zeros. In order for
this assumption to be realistic, we use a checkpoint close to the end
of the simulation so that most of the field is perturbed and hence is
close to the worst case in terms of compression results.

>>>>

Our performance model assumes
 that the compression algorithm behaves uniformly
across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
now becomes
\begin{equation}
\mathbf{O}_{SR} = \mathbf{N}_W \cdot \left(\frac{\mathbf{S}}{\mathbf{F}
  \cdot \mathbf{B}} + t_c\right) + \mathbf{N}_R \cdot
\left(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d\right)
\end{equation}
where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
and $t_d$ are compression and decompression times, respectively. At the same
time, the recomputation overhead decreases
because $\mathbf{F}$ times more checkpoints are now available.

@@

Figure~\ref{fig:varying_compute} shows the speedups predicted by the
performance model for kernels of different computational cost. Perhaps
unsurprisingly, the benefit gained by compression drops rapidly if the
time to compute a single timestep is less than the time to compress
the computed data (since recomputing would be cheaper in that case). For
increasing computational cost, the speedup from compression saturates
at a bound. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-compute.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  compute cost. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. The benefits of compression drop rapidly if the
computational cost of the kernel that generated the data is much lower
than the cost of compressing the data. For increasing computational
costs, the benefits are bounded.}
\label{fig:varying_compute}
\end{figure}

Figure~\ref{fig:varying_nt} shows the speedups predicted for differing
number of timesteps. It is interesting to note that there is a
consistent increase in the speedup as the number of timesteps goes
up. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-nt.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  number of timesteps to be reversed. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. It can be seen that compression becomes more
beneficial as the number of timesteps is increased.}
\label{fig:varying_nt}
\end{figure}

\section{Conclusions and Future work}
What are the problems?
The performance we developed might be useful in judging whether compression
would be helpful in a particular scenario without actually building a
complete implementation first. We have tried to keep it application
and platform agnostic in order to maximize its applicability. However
there are a few limitations in this model. Firstly, we assumed that
the compression performance we saw for a single checkpoint is
representative of any checkpoint throughout the simulation. This
assumption might have been completely wrong if made about a checkpoint
in the first few time steps since we consistently saw upto 800x
compression in those cases - since they are mostly zeros. In order for
this assumption to be realistic, we use a checkpoint close to the end
of the simulation so that most of the field is perturbed and hence is
close to the worst case in terms of compression results.
